### 2019-03-25 manuscript/manuscript.Rmd
Bryan's editorial suggestions and comments on the MS

---
title: "Long-term Water Quality Change in Rhode Island Lakes and Ponds"
author:
- Hollister. J. W. ^\*^ ^*1*^, Kellogg, D. Q. ^*2*^, Kreakie, B. J. ^*1*^, Shivers,
  S. ^*3*^, Milstead, W. Bryan. ^*1*^, Herron, E. ^*2*^, Green, L. ^*2*^, Gold, A
  ^*2*^
- ^*1*^ US Environmental Protection Agency, Office Of Research and Development, Atlantic
  Ecology Division, Narragansett, RI 02882
- ^*2*^ University of Rhode Island, Department of Natural Resources Science, Kingston,
  RI 02881
- ^*3*^ ORISE, Narragansett, RI 02882
- *'^\*^  corresponding author: hollister.jeff@epa.gov'*
bibliography: manuscript.bib
output:
  word_document:
    reference_docx: manuscript_template.docx
editor_options:
  chunk_output_type: inline
csl: plos.csl
abstract: The University of Rhode Island's Watershed Watch Volunteer Monitoring
  Program has been collecting water quality data on Rhode Island lakes and ponds for
  close to three decades, allowing us to explore long-term trends in common water
  quality parameters.  Not all lakes and ponds in the study area were sampled across
  the full time period and lakes were often added in geographic clusters (e.g. in
  urbanized northern Rhode Island).  Similar to the way long-term temperature records
  are analyzed, we centered and scaled (i.e., the z-score) water quality measurements
  on a per-station basis. This provides a robust and commonly scaled measurement to
  explore this data for long-term trends.  State-wide aggregation shows increasing
  temperature, chlorophyll *a*, and total nitrogen. Interestingly, total phosphorus
  is showing a decline, perhaps reflecting the management focus on phosphorus reductions.   While yearly trends are useful, they do mask month-to-month variability differences 
  across sites.  Additionally, while most sites track the yearly trend in decreasing water
  quality, there are bright spots with some sites improving over the 25 years. Contrary
  to previously reported analyses that show relatively stable water quality at the
  regional scale, our analysis shows that long-term water quality trends within Rhode
  Island show some parameters improving (which ones?) while others are in decline (which ones?).  Importantly,
  this analysis also points out the incredible value and importance of data from long-term
  monitoring programs, like the Univeristy of Rhode Island's Watershed Watch, for
  identifying trends in environmental condition.
---

```{r setup, include=FALSE}
source(here::here("R/functions.R"), encoding = "UTF-8")
opts_chunk$set(echo = TRUE, fig.width=6, fig.height=5)
tab_num <- captioner(prefix = "Table")
fig_num <- captioner(prefix = "Figure")
options(scipen = 2, digits = 5)
```

```{r analysis, include=FALSE}
#got to figure out path stuff before render.
source(here::here("R/data_available.R"), encoding = "UTF-8") 
source(here::here("R/summary_stats.R"), encoding = "UTF-8")  
source(here::here("R/figures.R"), encoding = "UTF-8")
```

# Introduction

Aquatic ecosystems have been altered as the result of human activities that modify bio-geochemical cycles on a global scale [@vitousek1997human; @finlay2013human; @filippelli2008global]. Because of their position in the landscape, lakes integrate and serve as sentinels for these anthropogenic effects [@williamson2008lakes; @schindler2009lakes]. Increasing nutrient inputs, particularly of N (NOTE: I think we should spell out nitrogen and phosphorus throughout-WBM) and P, derived from intensive agriculture and densely populated urban areas have contributed to the eutrophication of many lakes [@carpenter1998nonpoint; @smith2003eutrophication]. Eutrophication has caused the frequency and severity of harmful algal blooms (HABs) to increase, and these blooms are predicted to worsen under a warming climate [@michalak2013record; @kosten2012warmer; @taranu2015acceleration]. Increasing HABs will cause further declines in water quality, greater risks for human and animal health, and will be economically costly [@brooks2016harmful; @dodds2008eutrophication ;@paerl2009climate]. To address these problems, management strategies have historically focused on reducing P inputs to lakes, but research also suggests that concurrently reducing N inputs may be more effective in certain situations [@paerl2016takes; @schindler2008eutrophication]. These contrasting studies indicate that spatial differences and relationships between N, P, chlorophyll *a*, and HABs exist and that long-term studies are needed to identify trends at local, regional, and national scales.

Programs such as USEPA's National Lakes Assessment (NLA) provide data for continental scale water quality analysis. These analyses can be used for managing water resources by developing water quality criteria for N, P, chlorophyll *a*, and microcystin [@herlihy2013using; @yuan2014managing]. Studying trends across broad spatial scales can elucidate the effects of eutrophication such as the degradation of oligotrophic systems as P increases[@stoddard2016continental]. Continental scale data can also be used for water quality modeling across broad spatial scales including predicting lake trophic state and drawing inference on ecosystem condition [@hollister2016modeling; @nojavan2019rethinking]. These trophic state models indicate that landscape variables (i.e. ecoregion, elevation, and latitude) are important and that regional trends exist [ref???]. Lake-specific drivers are also important for predicting continental-scale water quality which adds an additional layer of complexity [@read2015importance]. Despite these challenges, it is important to study lake water quality at multiple spatial scales because emergent trends on regional or continental scales may not be evident when studying individual lakes [@cheruvelil2013multi; @lottig2014long].

Previous studies using regional data from the northeastern and midwestern United States have investigated spatial and temporal water quality trends and have shown differences based on scale. Macroscale (i.e. subcontinental) drivers of water quality trends are complex and may vary temporally [@lottig2017macroscale]. This complexity can cause nutrient (N and P) trends to have different drivers than ratios of the individual nutrients [@collins2017lake]. On a regional scale, trends of N, P, and chlorophyll *a* differ as factors such as land use and climate vary among regions, particularly when comparing the northeastern and midwestern US [@filstrup2014regional; @filstrup2018evidence]. Thus, it was surprising when stasis was reported over a 25 year period for these regions [@Oliver_2017]. Given what is known about long term trends in water quality within the broader region of the northeastern United States, we were curious if those trends were also present in water quality trends in Rhode Island lakes and ponds.

Examining long term trends in Rhode Island lakes is possible because of the data gathered by University of Rhode Island’s Watershed Watch (URIWW).   URI’s Watershed Watch is a XX year-old citizen science project that has built a robust collaboration between URI scientists and a vast network of volunteer monitors.  Over the past XX years, volunteer monitors have been trained to collect *in situ* data and whole water samples during the growing season.  These citizen science efforts allow for the collection of reliable data that can lead to crucial and frequently unexpected insights [@dickinson2012current; @kosmala2016assessing].  Here specifically, URIWW data contributed to not only the larger subcontinent extent study, but also allowed us to examine the long-term trends of Rhode Island.

Thus, the goals of this study were to look at approximately 25 years of lakes and ponds data in Rhode Island and answer two particular questions.  First, what are the state-wide trends in total nitrogen, total phosphorus, chlorophyll *a*, and lake temperature and second, are water quality trends in Rhode Island similar to regional trends in the northeastern United states.   In addition to these goals, this analysis has also been done using open data from the URI Watershed Watch program and the LAGOS project and the analysis in its entirety is available for independent reproduction at <https://github.com/usepa/ri_wq_trends> [NEED CITATIONS].

# Methods

For this study we combined a long-term dataset on water quality of lakes in Rhode Island, with a trend analysis based on centered and scaled water quality values (i.e. z-scores) to find increasing or decreasing annual water quality trends.  Details are outlined below.

## Study Area and Data

The study area includes all lakes and ponds in the state of Rhode Island monitored by the University of Rhode Island's Watershed Watch (URIWW: pehaps we'd be better off with "watershed watch" instead of an acronym-wbm) program (`r fig_num("ww_map", display = "cite", caption = "Map of URI Watershed Watch lake and pond sampling sites")`).  Watershed Watch is a volunteer monitoring program that has been collecting water quality data from Rhode Island lakes and ponds for over 30 years. The program began in 1988 with the monitoring of 14 lakes. The program has now grown to include greater than 250 monitoring sites, 120 waterbodies (including rivers/streams, and estuaries), and 400 trained volunteers (past and present?). Watershed Watch is an integral part of Rhode Idland's state environmental data collection strategy and provides more than 90% of the state's multi-year baseline data on lakes. Data QA/QC is of paramount importance; volunteers receive both classroom and field training and are provided with all the necessary equipment and supplies, along with scheduled collection dates. For freshwater lakes and ponds, weekly secchi depth and water temperature are recorded, along with bi-weekly Chl-a and dissolved oxygen. Water samples are collected three times per season (May through October) to be analyzed in the EPA-certified laboratory for nutrients, alkalinity, pH, and bacteria.

*`r fig_num("ww_map", caption = "Map of URI Watershed Watch lake and pond sampling sites: NOTE (wbm): since all figures have two symbols /colors why not use the same throughout and make sure each figure has a legend")`*
![](../figures/ww_map.jpg)

To understand the trends in lake temperature, total nitrogen, total phosphorus, and chlorophyll a we analyzed a subset of the watershed watch data.  We chose sample sites with at least 10 years of data that were sampled at a depth of 2 m or less during the months of May to October from 1990 to 2016. The final dataset included `r nrow(ww_avail_data_locations)` lakes and ponds `r sum(ww_avail_data_locations$params_avail == "chla, total_p, total_n, temp")` of which had complete samples of all 4 parameters.

To place our results in context we reanalyzed the data from prior studies that looked at water quality trends across a larger region of the Northeastern and Midwestern United states that included 17 states (Minnesota, Wisconsin, Iowa, Missouri, Illinois, Indiana, Michigan, Ohio, Pennsylvania, New York, New Jersey, Connecticut, Massachusetts, Rhode Island, Vermont, New Hampshire, and Maine) [@soranno2015building; @Oliver_2017].  The authors found little change in water quality trends across this region [@Oliver_2017] .  We repeated our analysis (see **Trend Analysis** section) with the same dataset used by [@Oliver_2017], the LAGOSNE dataset [@soranno2015building; @stacheleklagosne2017; @sorannolagosne2017].  Temperature data were not availble, thus we only examine trends, using our analytical methods, for Total Nitrogren, Total Phosphorus, and Chlorophyll *a* from the LAGOSNE database.

## Trend Analysis

There are many different methods for analyzing time series data for trends.  One of the difficulties that is encountered when you have multiple sampling locations is how to identify a trend when you have spatial and temporal variation in sampling effort.  For instance, if you have long-term data on water quality for several ponds, yet the cleaner ponds wer sampled more frequently in early years then a simple comparison of raw-values over time might show a decrease in water quality.  Thus, it is necessary to account for this type of site-specific variation.  This is similar to how the long-term temperature trends are analyzed using temperature anomolies [e.g., https://www.ncdc.noaa.gov/teleconnections/enso/indicators/sst/].  The general approach is to calculate site-specific deviations from a long-term mean over pre-determined reference period.  This allows all sites to be shifted to a common base line and the deviations, or anomalies, show change relative to the reference period.

NOTE(WBM): There is also local variation in history, lake morphometry, and ecology that can affect the data.  Scaling and centering removes these confounding effects by converting all values to a mean of zero and a sd of 1.

### Z-score

The anomaly approach is very useful and is the approach we chose to follow.  However, using anomalies works best with a single measure (e.g temperature) or with mulitple measurements that are on the same scale.  The water quality parameters that we explored have different scales and thus the anomaly alone is difficult to interpret across metrics.  For instance, temperature in Rhode Island lakes during the growing season will range from approximately 15 degrees celsius to a high of 30 degrees celsius, whereas phosphorus might range from near zero ug/l to ~900 ug/l.  To standardize these values we used the common approach of dividing each anomlaly by the standard deviation for the reference period.  This value is commonly referred to as a Z-score.  We use these z-scores to examine each water quality variable for a trend over the time period of 1990 to 2016.

Furthermore, since we are interest in water quality trends over time, we wanted to explore how each site was responding.  Thus, our z-scores were calculated over the reference period, 1990-2016, for each site.  In other words the long-term mean and standard deviation were calculated on a site-by-site basis. NOTE(wbm): So the index period will vary among lakes.  I can image some reviewers will balk at this.  Can we justify?

### Linear regression

We used linear regression of z scores as a function of year to test for monotonic trends in the lake water quality data [@helsel2002statistical].  Simply stated, this tests the null hypothesis that no trend exists (e.g. β1 = 0). If the null hypothesis is rejected a slope greater than zero indicates a positive or rising trend while a negative slope shows a negative or decreasing trend.  Rejection of the null hypothesis for  lake temperature, chlorophyll *a*, total nitrogen, and / or total phosphorus indicates positive or negative trends for these parameters (depending on the sign of the slope).  For each parameter we fit a regression line to the z scores as a function of year and tested the null hypothesis that no trend existed (e.g. $\beta$~1~ = 0).

### Comparison of Rhode Island to the Region

Oliver et al. [@Oliver_2017] examined a larger dataset for the lakes of the Northeastern and Midwestern United States and found relatively stable water quality in the lakes over time.  The University of Rhode Island's Watershed Watch data were inlcuded in this larger study. To assess whether regional trends were masking local scale trends in Rhode Island we re-analyzed the Oliver et al. [@Oliver_2017] dataset using the approach outlined above.

# Results

During the sampling period of 1990 to 2016 Rhode Island lakes and ponds average lake temperature was `r paste(ww_param_ranges%>%filter(Parameter == "temp")%>%pull(Mean), ww_param_ranges%>%filter(Parameter == "temp")%>%pull(Units))`, average total nitrogen was `r paste(ww_param_ranges%>%filter(Parameter == "total_n")%>%pull(Mean), ww_param_ranges%>%filter(Parameter == "total_n")%>%pull(Units))`, average total phosphorus was `r paste(ww_param_ranges%>%filter(Parameter == "total_p")%>%pull(Mean), ww_param_ranges%>%filter(Parameter == "total_p")%>%pull(Units))`, and average chlorophyll was `r paste(ww_param_ranges%>%filter(Parameter == "chla")%>%pull(Mean), ww_param_ranges%>%filter(Parameter == "chla")%>%pull(Units))` (`r tab_num("ww_params", display = "cite", caption = "Summary statistics for URI Watershed Watch data from 1990 to 2016.")`).

*`r tab_num("ww_params")`*

```{r ww_params, echo=FALSE, results="asis"}
knitr::kable(ww_param_ranges, format = "markdown")
```

For lakes in ponds in the larger region represented by the LAGOSNE States,  average total nitrogen was `r paste(lagos_param_ranges%>%filter(Parameter == "total_n")%>%pull(Mean), lagos_param_ranges%>%filter(Parameter == "total_n")%>%pull(Units))`, average total phosphorus was `r paste(lagos_param_ranges%>%filter(Parameter == "total_p")%>%pull(Mean), lagos_param_ranges%>%filter(Parameter == "total_p")%>%pull(Units))`, and average chlorophyll was `r paste(lagos_param_ranges%>%filter(Parameter == "chla")%>%pull(Mean), lagos_param_ranges%>%filter(Parameter == "chla")%>%pull(Units))` (`r tab_num("lagos_params", display = "cite", caption = "Summary statistics for LAGOSNE data from 1990 to 2016.")`).

*`r tab_num("lagos_params")`*

```{r lagos_params, echo=FALSE, results="asis"}
knitr::kable(lagos_param_ranges, format = "markdown")
```

## State-wide trends in water quality

Average yearly scaled temperature in lakes and ponds appear to be increasing (slope: `r round(ww_temp_rl$slope,3)` , p-value: `r round(ww_temp_rl$p.value, 5)`) with the large majority of years with average temperature greater than the long term average occuring in the years since 2000 (`r fig_num(name = "temp_ri", caption = "Yearly trend over 20+ years of lake temperature in Rhode Island lakes and ponds.", display = "cite")`).  Chlorophyll *a* is also showing an increasing trend over time (slope: `r round(ww_chla_rl$slope,3)` , p-value: `r round(ww_chla_rl$p.value, 5)`) and with the exception of an early high chlorophyll *a* year in 1992 and a slightly above average year in 2003, the above average years have all occurred since 2010 (`r fig_num(name = "chla_ri", caption = "Yearly trend over 20+ years of Chlorphyll a (average z-score) in Rhode Island lakes and ponds.", display = "cite")`).

*`r fig_num(name = "temp_ri")`* In this and subsequent figures values above or below the long term average are shown in red and blue respectively (NOTE: we should also use different symbols).

![](../figures/ww_temp_trends.jpg)

*`r fig_num(name = "chla_ri")`*

![](../figures/ww_chla_trends.jpg)

```{r no_early_tn, echo=FALSE}
ww_tn_rl_no9394 <- lm(mn_value~year, data = ww_tn_df[c(-1,-2),]) %>%
  tidy() %>%
  slice(2) %>%
  select(slope = estimate, p.value)
```

Average yearly trends for nutrients showed weaker or no trends over time.  Total nitrogen did have a significant (NOTE(wbm): Still, it looks like there could be a trend.  Should we run the regression without the 1993 and 1994 data?) trend (slope: `r round(ww_tn_rl$slope,3)` , p-value: `r round(ww_tn_rl$p.value, 5)`); however, that trend is driven by the lower than average total nitrogen values in 1993 and 1994 (`r fig_num(name = "tn_ri", caption = "Yearly trend over 20+ years of Total Nitrogen (average z-score) in Rhode Island lakes and ponds.", display = "cite")`).  Since 1995, the yearly trend is not significant (slope: `r round(ww_tn_rl_no9394$slope,3)`, p-value: `r round(ww_tn_rl_no9394$p.value, 5)`).  Total phosphorus does not show a trend over time in the yearly scaled values (slope: `r round(ww_tn_rl$slope,3)` , p-value: `r round(ww_tn_rl$p.value, 5)`) and years that are over or under the average are evenly distributed over the years (`r fig_num(name = "tp_ri", caption = "Yearly trend over 20+ years of Total Phosphorus (average z-score) in Rhode Island lakes and ponds.", display = "cite")`).

*`r fig_num(name = "tn_ri")`*

![](../figures/ww_tn_trends.jpg)

*`r fig_num(name = "tp_ri")`*

![](../figures/ww_tp_trends.jpg)

## Regional trends in water quality

In general, broad regional trends were less obvious and in the case of chlorophyll *a* and total phosphorus, the trends were not significant.  In particular, chlorophyll *a* showed a slight, yet non-significant negative trend (slope: `r round(lagos_chla_rl$slope,3)`, p-value: `r round(lagos_chla_rl$p.value, 5)`, `r fig_num(name = "chla_lagos", caption = "Yearly trend over 20+ years of Chlorophyll a (average z-score) in the LAGOSNE states.", display = "cite")`).

*`r fig_num(name = "chla_lagos")`*

![](../figures/lagos_chla_trends.jpg)

Total phosphorus shows a slight and non-significant increasing trend (slope: `r round(lagos_tp_rl$slope, 3)`, p-value: `r round(lagos_tp_rl$p.value,4)`, `r fig_num(name = "tp_lagos", caption = "Yearly trend over 20+ years of Total Phosphorus (average z-score) in the LAGOSNE states.",display = "cite")`) and total nitrogen shows a slight negative non-significant trend (slope: `r round(lagos_tn_rl$slope, 3)`, p-value: `r round(lagos_tn_rl$p.value, 5)`, `r fig_num(name = "tn_lagos", caption = "Yearly trend over 20+ years of Total Nitrogen (average z-score) in the LAGOSNE states.", display = "cite")`).

*`r fig_num(name = "tp_lagos")`*

![](../figures/lagos_tp_trends.jpg)

*`r fig_num(name = "tn_lagos")`*

![](../figures/lagos_tn_trends.jpg)

# Discussion and conclusions

## Volunteer Monitoring Data

### Why this is possible - Fantastic WW data

## Trends

### Parameter by parameter discussion of trends

### How do trends break down by trophic state?

### Local scale vs Regional/Global scale

### RI is downstream - Bryan will write

## Management/Implications

### Season Length Changes

## Caveats

# Bibliography